---
title: "Report for Reproducible Research - Assessment1"
author: "Roberto Lo Prejato"
date: "25 ottobre 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In thi assignment I make use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

```{r, results='hide', warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(lubridate)
```
## Loading and preprocessing the data

First of all I need to load the file.csv with data and store it in a variable names "activity""
```{r load data}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity <- read.csv("activity.csv")
```

In the following histogram we can see the total numer of steps taken per day 
```{r histogram}
act_na <- filter(activity, !is.na(steps))
activity_sum <- act_na %>% group_by(date) %>% summarize(st_tot= sum(steps))
qplot(st_tot, data=activity_sum, bins=30, main="Total number of steps per day", xlab="steps")
```

```{r report mean and median, results='hide'}
mean_steps <- round(mean(activity_sum$st_tot), digits = 2)
median_steps <- median(activity_sum$st_tot)
```
The mean the median of steps per day is respectively `r mean_steps` and `r median_steps`. 

Considering the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) we can see the average daily activity pattern.
```{r time series}
activity_interval <- act_na %>% group_by(interval) %>% summarize(st_mean= mean(steps))
interval_max <- activity_interval%>%filter(st_mean == max(st_mean))%>%select(1)
p1 <- ggplot(activity_interval, aes(x = interval , y = st_mean))
p1+geom_line(color="blue", size=1) + labs(title = "Avg. Daily Steps", x = "Interval", y = "Avg. Steps per day")
```
The 5-minute interval with most steps on average across all the days is `r interval_max`.

##Imputing missing values
```{r missing value, results='hide'}
row_na <- nrow(activity)-nrow(act_na)
row_na <- nrow(activity)-nrow(act_na)
x<-activity %>% group_by(date) %>% summarize(tot_steps = sum(steps)) %>% filter(is.na(tot_steps))
day_na <- nrow(x)
```
In this database there are `r row_na` row with null value as steps and `r day_na` dayes with no data for steps.
Now we will fill the NA value for `r row_na` row with the median of steps fot the specific 5-minutes interval so we can create a new database without null value in steps column
```{r new dataset, results='hide'}
data_fill <- act_na %>% group_by(interval) %>% summarize(st_median = median(steps))
act_fill <- activity %>% filter(is.na(steps)) %>% merge(data_fill) %>% select(-steps) %>% rename(steps =st_median)
#create a new db filling NA value
activity_fill <- bind_rows(act_fill,act_na)

```

With this new dataset we can make again the histogram with the the total numer of steps taken per day and chack if somethings will change. 

```{r histogram2}
activity_sum2 <- activity_fill %>% group_by(date) %>% summarize(st_tot= sum(steps))
require(gridExtra)
plot1<-qplot(st_tot, data=activity_sum, bins=30, main="Steps per day - original", xlab="steps")
plot2<-qplot(st_tot, data=activity_sum2, bins=30, main="Steps per day - filled", xlab="steps")
grid.arrange(plot1, plot2, ncol=2)

mean_steps2 <- round(mean(activity_sum2$st_tot), digits = 2)
median_steps2 <- median(activity_sum2$st_tot)
```
Look at new median and mean:
-mean total number of steps taken per day = `r mean_steps`
-median total number of steps taken per day = `r median_steps`
We can see that there are some differences using the median do fill the null valune of steps, also the mean and median appears to be affected by this simple data imputation, theis value are less.

##Are there differences in activity patterns between weekdays and weekends?

We use the dataset with the filled-in missing values for this part.
we want to see if there is a different behavior between the weekdays the weekend.
For this reason we create a new factor variable in the dataset with two levels ("Weekday" and "Weekend") indicating whether a given date is a weekday or weekend day.

```{r week panel plot}
activity_fill$dayofweek <- weekdays(as.Date(activity_fill$date))
activity_fill$weekend <-as.factor(activity_fill$dayofweek=="sabato"|activity_fill$dayofweek=="domenica")
levels(activity_fill$weekend) <- c("Weekday", "Weekend")
act_week <- activity_fill %>% group_by(interval, weekend) %>% summarize(st_mean= mean(steps))
ggplot(act_week, aes(x = interval , y = st_mean)) + geom_line(color="blue", size=1) + labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + facet_wrap(~weekend , ncol = 1, nrow=2)
```

From the graph we can see that during the weekend the number of steps is higher
in the middle of the day

